---
title: "PostFireTrajectories and Model Prep for plot data"
author: "Jasper Slingsby, Adam M. Wilson and Glenn Moncrieff"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output:
  html_document:
    keep_md: yes
    number_sections: yes
    theme: cerulean
    toc: yes
  pdf_document:
    toc: yes
---

```{r,setup,echo=F,results='hide',warning=FALSE, message=FALSE, cache = F}
##  First some set up
source("/Users/glennmoncrieff/Documents/Projects/postfire/postfire/workflow/setup.R")
ig=raster(paste0(datadir,"clean/indexgrid_modis_250m.grd")) 
```


# Data

Load the model data we made in DataPrepFullRecord_plots.R
```{r loaddata,results='asis',warning=FALSE, message=FALSE, cache = F}
load("data/modeldata_dim_fix.Rdata")

plotdat <- left_join(tdat,sdat,by="plot")

#site we want to keep
plotsites <- c("CP","LB1","LB2","CB","BK")
dat <- filter(plotdat,Site.y%in%plotsites)%>% #filter site
                filter(QA==0)%>%filter(ND>0) %>%filter(complete.cases(.)) #filter low quality and nodata

kable(head(dat),row.names=F)
```


## Change through time

lets look at all plots by age - up to 10 year old plots

```{r,htmlcap="NDVI vs age for all plots",warning=FALSE, message=FALSE, cache = F}
P <- ggplot(dat,aes(x=DA, y=ND,color=Site.y,group=plot)) +
geom_line(size=0.05) +
coord_cartesian(ylim=c(0,1),xlim=c(0,3650)) +
#scale_color_manual(values=c("grey60")) +
facet_wrap( ~ region, ncol=2) +
labs(x = "days since fire", y = "NDVI") +
scale_colour_discrete(guide = FALSE) +
theme_bw() 
 
suppressWarnings(print(P))
```

It's still hard to see change while looking at the entire dataset, so let's:

1. pick a few plots and plot NDVI as a function of age  
```{r,htmlcap="NDVI vs age",warning=FALSE, message=FALSE, cache = F}
dat_plotname <- c("LB1_29","LB1_30","LB2_50","LB2_51","CP_45","CP_49","CB_66","CB_69","BK_38","BK_58")

dat_plot <- filter(dat,plot_smpl%in%dat_plotname)

P <- ggplot(dat_plot,aes(x=DA, y=ND,color=Site.y,group=plot)) +
geom_line(size=0.2) +
coord_cartesian(ylim=c(0,1),xlim=c(0,3650)) +
#scale_color_manual(values=c("grey60")) +
facet_wrap( ~ plot_smpl, ncol=3) +
labs(x = "days since fire", y = "NDVI") +
scale_colour_discrete(guide = FALSE) +
theme_bw() 
 
suppressWarnings(print(P))
```

2.  plot NDVI as a function of date: 
```{r,htmlcap="NDVI vs date",warning=FALSE, message=FALSE, cache = F}

P <- ggplot(dat_plot,aes(x=date, y=ND,color=Site.y,group=plot)) +
geom_line(size=0.2) +
scale_x_date()+
coord_cartesian(ylim=c(0,1)) +
#scale_color_manual(values=c("grey60")) +
facet_wrap( ~ plot_smpl, ncol=3) +
labs(x = "date", y = "NDVI") +
scale_colour_discrete(guide = FALSE) +
theme_bw() 
 
suppressWarnings(print(P))
```

# Non-linear model fitting

The full model I've been using (minus the seasonal component) says that the expected NDVI at some location $i$ in time $t$ comes from a normal distribution as follows:

$\text{NDVI}_{i,t}\sim\mathcal{N}(\mu_{i,t},\sigma)$ 

where the mean ($\mu$) is a nonlinear function including the post-fire NDVI value ($\alpha$), the potential increase in NDVI ($\gamma$), and the post-fire recovery rate ($\lambda$) as follows:

$\mu_{i,t}=\alpha_i+\gamma_i\Big(1-e^{-\frac{age_{i,t}}{\lambda_i}}\Big)$

#fit model

 Let's see what happens when we fit all pixels at once.

```{r fitmodel1,results='hide',warning=FALSE, message=FALSE, cache = F}
sform=as.formula(ND~alpha+gamma*(1-exp(-DA/lambda)))
m <- nlsLM(sform, data =dat, start = list(lambda=0.1,alpha=0.1,gamma=0.1), trace = T)
summary(m)
```

Plot it:

```{r fitmodel2,warning=FALSE, message=FALSE, cache = F}
## make a new dataframe of predictions from the model
x=seq(1:3650)
dpred=data.frame(pred=predict(m,newdata=list(DA=x)),x=x)

P <- ggplot()+
geom_line(data=dat,aes(x=DA,y=ND,group=plot),size=0.4,alpha=0.05) +
geom_line(data=dpred,aes(y=pred,x=x),colour="red",size=1) +
coord_cartesian(ylim=c(0,1),xlim=c(0,3650)) +
#scale_color_manual(values=c("grey60")) +
labs(x = "days since fire", y = "NDVI") +
scale_colour_discrete(guide = FALSE) +
theme_bw() 
 
suppressWarnings(print(P))

```

fit separate model for each site:


```{r fitmodel1sep,warning=FALSE, message=FALSE, cache = F}
## make a new dataframe of predictions from the model

CP <- filter(dat,Site.y=="CP")
mCP <- nlsLM(sform, data =CP, start = list(lambda=400,alpha=0.4,gamma=0.1), trace = T)

LB1<- filter(dat,Site.y=="LB1")
mLB1 <- nlsLM(sform, data =LB1, start = list(lambda=400,alpha=0.4,gamma=0.1), trace = T)

LB2<- filter(dat,Site.y=="LB2")
mLB2 <- nlsLM(sform, data =LB2, start = list(lambda=400,alpha=0.4,gamma=0.1), trace = T)

CB<- filter(dat,Site.y=="CB")
mCB <- nlsLM(sform, data =CB, start = list(lambda=400,alpha=0.4,gamma=0.1), trace = T)

BK<- filter(dat,Site.y=="BK")
mBK <- nlsLM(sform, data =BK, start = list(lambda=400,alpha=0.4,gamma=0.1), trace = T)

x=seq(1:3650)

dpredCP=data.frame(pred=predict(mCP,newdata=list(DA=x)),x=x,lab="CP")
dpredLB1=data.frame(pred=predict(mLB1,newdata=list(DA=x)),x=x,lab="LB1")
dpredLB2=data.frame(pred=predict(mLB2,newdata=list(DA=x)),x=x,lab="LB2")
dpredCB=data.frame(pred=predict(mCB,newdata=list(DA=x)),x=x,lab="CB")
dpredBK=data.frame(pred=predict(mBK,newdata=list(DA=x)),x=x,lab="BK")

```

Plot it:

```{r fitmodel2sep,warning=FALSE, message=FALSE, cache = F}

P <- ggplot()+
geom_line(data=dat,aes(x=DA,y=ND,group=plot),size=0.4,alpha=0.05) +
geom_line(data=dpredCP,aes(y=pred,x=x,colour=lab),size=1) +
geom_line(data=dpredLB1,aes(y=pred,x=x,colour=lab),size=1) +
geom_line(data=dpredLB2,aes(y=pred,x=x,colour=lab),size=1) +
geom_line(data=dpredCB,aes(y=pred,x=x,colour=lab),size=1) +
geom_line(data=dpredBK,aes(y=pred,x=x,colour=lab),size=1) +
coord_cartesian(ylim=c(0,1),xlim=c(0,3650)) +
#scale_color_manual(values=c("grey60")) +
labs(x = "days since fire", y = "NDVI") +
#scale_colour_discrete(guide = FALSE) +
theme_bw() 
 
suppressWarnings(print(P))

```


```{r,purl,eval=FALSE,echo=FALSE,results='hide',messages=F,error=FALSE}
## this chunk outputs a copy of this script converted to a 'normal' R file with comments
purl("workflow/3_PostFireTrajectories/PostFireTrajectories.Rmd",
     documentation=2,output = "workflow/3_PostFireTrajectories/PostFireTrajectories.R", quiet = TRUE) 
```
