---
title: "Modeling PostFireTrajectories"
author: "Adam M. Wilson"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output:
  html_document:
  keep_md: yes
number_sections: yes
theme: cerulean
toc: yes
pdf_document:
  toc: yes
---

```{r,setup,echo=F,cache=T,results='hide',message=FALSE}
##  First some set up
source("../setup.R")
ig=raster(paste0(datadir,"clean/indexgrid_landsat_30m.grd")) 
```


# Data

Load the model data we made in [DataPrep.R](../1_Data/DataPrep.R)
```{r loaddata,results='asis'}
load("data/modeldata.Rdata")
rv_meta=read.csv("data/vegtypecodes.csv")
sdat$vegn=rv_meta$code[match(sdat$veg,rv_meta$ID)]
rownames(sdat)=sdat$id
```

We know have two data frames we'll use in the modeling. The first is the spatial data:
```{r sdat, results='asis'}
kable(head(sdat),row.names=F)
```

And the second is the temporal data:
```{r tdat, results='asis'}
kable(head(tdat),row.names=F)
```

#dat=cbind.data.frame(tdat,sdat[match(tdat$id,sdat$id),])


```{r CleanData}

#ndvi=ndvi[!ndvi$id%in%names(tapply(ndvi$agemonth,ndvi$id,function(x) max(x)<36))[tapply(ndvi$agemonth,ndvi$id,function(x) max(x)<36)],] #drop records with no data less than 3 years old
#ndvi=ndvi[!ndvi$id%in%names(table(ndvi$id))[table(ndvi$id)<2*12*3],] #drop records with less than 3 years of data

## Keep only fynbos veg types
sdat=sdat[grepl("Fynbos",sdat$vegn),]

#drop obs of unknown age or ndvi or with no data in sdat
tdat=tdat[!is.na(tdat$age)&!is.na(tdat$ndvi)&tdat$id%in%sdat$id,] 

## drop negative ages (time before first fire) for now
tdat=tdat[tdat$age>=0,]

########################
#### Set model name for naming objects below and create directory to hold output
mname="v1"
if(!file.exists(paste("output/",mname,sep=""))) dir.create(paste("output/",mname,sep=""),recursive=T)

### subset dataset
holdout=0.90  #percent to hold out for validation
s=sort(sample(unique(sdat$id),round(length(unique(sdat$id))*(1-holdout)))); length(s)
write.csv(s,paste("output/",mname,"/",mname,"_subset.csv",sep=""),row.names=F)
sdat$subset=factor(ifelse(sdat$id%in%unique(tdat$id),ifelse(sdat$id%in%s,"Model Fitting","Validation"),"Prediction"),levels=c("Model Fitting","Validation","Prediction"),ordered=T)
```

## Create dummy variables for vegetation (and any other factors)
```{r dummy,results='asis'}
sdat$veg=as.factor(sdat$veg)
lm1=lm(dem~veg,data=sdat)
tveg=model.matrix(lm1)[,-1]

kable(head(tveg))
```

```{r scale}
## Select and scale environmental data
envars=c("dem","tpi","tmax","tmin")

scaled=scale(as.matrix(sdat[,envars]))
env_full=cbind(intercept=1,scaled,tveg)

### Save the scaling parameters to convert fitted coefficients back to metric units later
beta.mu=c(intercept=0,attr(scaled,"scaled:center"),veg11=0,veg14=0,veg15=0,veg16=0,veg17=0)
beta.sd=c(intercept=1,attr(scaled,"scaled:scale"),veg11=1,veg14=1,veg15=1,veg16=1,veg17=1)
rm(scaled)  #drop the scaled data
```


## Set up the data for modeling
```{r modeldata}
tdat_full=tdat
tdat=tdat[tdat$id%in%s,]; gc() 

## create two env frames for fitting and prediction
env=env_full[rownames(env_full)%in%s,]
  
### Drop missing values
omit=unique(tdat$id)[as.numeric(which(is.na(apply(env,1,sum))))]; omit
if(length(omit)>0){
  env=env[!rownames(env)%in%omit,]
  tdat=tdat[!tdat$id%in%omit,]
}

## create new id that goes from 1 to nGrid
tdat$id2=as.integer(as.factor(tdat$id)); gc()

## Get counts
nGrid=length(unique(tdat$id))            ;nGrid
nTime=length(unique(tdat$year))          ;nTime
nBeta=ncol(env)                          ;nBeta

## Write data object
data=list(
  age=tdat$age,
  ndvi=tdat$ndvi,
  id=tdat$id2,
  nObs=nrow(tdat),
  env=env,
  nGrid=nGrid,
  nBeta=nBeta
  )

## Function to generate initial values
gen.inits=function(nGrid,nBeta) { list(
  ## spatial terms
  alpha=runif(nGrid,0.1,0.5),
  gamma=runif(nGrid,0.1,.9),
  lambda=runif(nGrid,0.2,1),
  ## spatial means
  alpha.mu=runif(1,0.1,0.2),
  ## priors  
  gamma.beta=runif(nBeta,0,1),
  gamma.tau=runif(1,1,5),
  alpha.tau=runif(1,1,5),
  lambda.beta=runif(nBeta,0,2),
  lambda.tau=runif(1,0,2),
  tau=runif(1,0,2)
  )
}

## list of parameters to monitor (save)
params=c("gamma.beta","gamma.sigma","alpha","alpha.mu","alpha.sigma","lambda.beta","lambda.sigma")

### Save all data into Rdata object for model fitting
save(data,gen.inits,s,sdat,beta.mu,beta.sd,envars,env_full,tdat_full,file=paste("output/",mname,"/",mname,"_inputdata.Rdata",sep=""))  #save it
```


# JAGS
```{r jags}

library(rjags)

write.table(paste("Starting model ",mname," on ",date()),paste0("output/",mname,"ModelStart.txt"))

## test compilation
t1=system.time(m <<- jags.model(file="workflow/6_Modeling/Model.R",
                             data=data,
                             inits=gen.inits(data$nGrid,data$nBeta),
                             n.chains=1,n.adapt=100))
```

### Collect 'Posterior Samples' and DIC
```{r}
t2=system.time(mc <<- coda.samples(m,params,n.iter=100))
#dic=dic.samples(m,n.iter=10,type="pD")
```

### Save output
```{r}
save(m,mc,beta.mu,beta.sd,
     file=paste("output/",mname,"/",mname,"_modeloutput.Rdata",sep=""))  
```

