---
title: "PostFireTrajectories"
author: "Adam M. Wilson"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output:
  html_document:
    keep_md: yes
    number_sections: yes
    theme: cerulean
    toc: yes
  pdf_document:
    toc: yes
---

```{r,setup,echo=F,cache=F,results='hide',message=FALSE}
##  First some set up
source("../setup.R")
ig=raster(paste0(datadir,"clean/indexgrid_landsat_30m.grd")) 
```


# Data

Load the model data we made in [DataPrep.R](../3_DataAnnual/DataPrep.Rmd)
```{r loaddata,results='asis'}
load("data/modeldata_annual.Rdata")
rv_meta=read.csv("data/vegtypecodes.csv")
sdat$vegn=rv_meta$code[match(sdat$veg,rv_meta$ID)]

## now create a single monster table with all the data
dat=cbind.data.frame(tdat,sdat[match(tdat$id,sdat$id),])
## drop negative ages (time before first fire) for now
dat=dat[dat$age>=0,]
## look at the table
kable(head(dat),row.names=F)
```


## Change through time

It's still hard to see change while looking at the full peninsula, so let's:

1. pick a few pixels and plot NDVI as a function of time.  
2. zoom in on a smaller region

Let's load the NDVI data again.
```{r loadndvi}
## load the NDVI data
ndvifile="data/ndvi_annual_landsat_30m.tif"
years=1984:2014
ndvi=stack(ndvifile)
names(ndvi)=paste0("ndvi_",years)
ndvi=setZ(ndvi,years)
```

But first we need to pick a pixel (or a few) to plot.  There are a few ways to do this.  First let's try extracting values from a few cells:
```{r pickpoints}
## first plot the data
plot(ndvi[[1]])
## select a few points by row number
d=data.frame(cell=c(1095098,1070102,1006689))
points(sdat[match(d$cell,sdat$id),c("x","y")],pch=16)
text(sdat[match(d$cell,sdat$id),c("x","y")],labels=d$cell,pos=4)
```

To select points for plotting, we can use the `click` function (in Raster) that allows you to pick points on the map.  First we need to plot the data using the plot() command. Then run the `click` command using the `ig` grid so the cell id's are returned. 
```{r pickpoints2,eval=FALSE}
## first plot the data
plot(ndvi[[2]])
## specify how many points you want to select
nclicks=5
## run the click function then click on the map (to the right).  This may not work in GUI's other than RStudio...
d=click(ig,n=nclicks,cell=T,id=T)
```


Now plot that as a function of time (year).
```{r,htmlcap="NDVI vs year for a selection of pixels"}
ggplot(dat[dat$id%in%d$cell,],
       aes(x=year,y=ndvi,group=id))+
  geom_line(aes(color = factor(id)),size=1,linetype=id)
```

And as a function of age
```{r,htmlcap="NDVI vs age for a selection of pixels"}
ggplot(dat[dat$id%in%d$cell,],
       aes(x=age,y=ndvi,group=id))+
  geom_line(aes(color = factor(id)),size=1)+
  stat_smooth(fill = "grey50",aes(group = 1),col="red")
```

  Explore the NDVI data for various pixels by changing `nclicks` above and selecting new points on the map.  Remember that we subsetted the dat dataframe to include only 'natural vegetation' so you may select a point with no data.  

### Regional plot

Alternatively, we can aggregate the data from a larger region.  First import a shapefile of the reserves on the peninsula and the fire data.
```{r loadreserves,results='hide'}
## load the reserve shapefile
reserves=readOGR(paste0(datadir,"raw/reserves/"),"reserves")
reserves=spTransform(reserves,CRS(proj4string(ig)))

## or pick a fire polygon
fi=readOGR(dsn=paste0(datadir,"raw/Fire"), layer="CapePenFires") #Cape Peninsula fires history layers 1962-2007
## transform to working projection
fi=spTransform(fi,CRS(proj4string(ig)))
```

Now select a region to explore.  You could do a single fire, a single reserve, or any combination of regions using the code below.  
```{r subset,htmlcap="Elevation in selected region"}
## select a reserve
resname="SILVERMINE"
reg1=reserves[which(reserves$MASTERNAME==resname),]

## or pick a fire
#reg1=fi[which(fi$FIREID==2000103),]
## get cell numbers in that region

## Extract the data for that polygon
rd=extract(ig,reg1)[[1]]


ggplot(sdat[sdat$id%in%rd,], aes(x=x,y=y))+
   geom_tile(aes(fill=dem))

```

Let's look at all those pixels through time:
```{r regextract,results='hide',warning=FALSE}
ggplot(dat[dat$id%in%rd,],aes(x=as.numeric(year),y=ndvi,group=id))+
  geom_line(size=.2,alpha=.1)+
  stat_smooth(fill = "grey50",aes(group = 1),col="red")+
  coord_fixed(ratio = 80)
```

And vs. cell age
```{r,results='hide',warning=FALSE}
ggplot(dat[dat$age>=0&dat$id%in%rd,],aes(x=age,y=ndvi,group=id))+
  geom_line(size=.2,alpha=.1)+facet_wrap(~vegn,nrow=1)+
  stat_smooth(fill = "grey50",aes(group = 1),col="red")#+xlim(0, 30)
```

# Non-linear model fitting

The full model I've been using (minus the seasonal component) says that the expected NDVI at some location $i$ in time $t$ comes from a normal distribution as follows:

$\text{NDVI}_{i,t}\sim\mathcal{N}(\mu_{i,t},\sigma)$ 

where the mean ($\mu$) is a nonlinear function including the post-fire NDVI value ($\alpha$), the potential increase in NDVI ($\gamma$), and the post-fire recovery rate ($\lambda$) as follows:

$\mu_{i,t}=\alpha_i+\gamma_i\Big(1-e^{-\frac{age_{i,t}}{\lambda_i}}\Big)$

## Create a subset
To make things run faster while we're experimenting, let's subset the data,  For example, drop ages before the first fire (which are censored and uncertain), keep only fires in `r resname`, and select only "Peninsula Sandstone Fynbos"


Let's look at that timeseries for all points in the subsetted region.  
```{r p1}
dats=dat[dat$vegn=="Peninsula Sandstone Fynbos"&dat$id%in%rd,] 
dats=dats[!is.na(dats$id),]

ggplot(dats,aes(x=age,y=ndvi,group=id))+
  geom_line(size=.1,alpha=.1)
```

Woah, that's messy.  Could there be any information there?  Let's see what happens when we fit all pixels at once.

```{r fitmodel1,results='hide'}
## Assign starting values for all parameters.  The search has to start somewhere...
start=list(alpha=0.2,gamma=0.4,lambda=4)
## define lower and upper bounds for the parameters to limit the search
lower=c(0,0,0)
upper=c(1,1,10)
## other nls control settings
ctl=nls.control(maxiter = 150,minFactor=1e-10)
## Assign ages at which to evaluate the model
x=seq(0,30,len=100)


sform=as.formula(ndvi~alpha+gamma*(1-exp(-age/lambda)))
m <- nlsLM(sform, data =dats, start = start, trace = T,control=ctl,lower=lower,upper=upper)
summary(m)
```

Plot it:

```{r fitmodel2}
## make a new dataframe of predictions from the model
dpred=cbind.data.frame(ndvi=predict(m,newdata=data.frame(age=x)),age=x,id=1)

ggplot(dats,aes(x=age,y=ndvi,group=id))+
  geom_line(size=.2,alpha=.2)+
  geom_line(data=dpred,aes(y=ndvi,x=age),colour="red",size=2)

```

Do you believe it?  Useful?  How to improve upon this approach?  What other factors are important?



```{r,purl,eval=FALSE,echo=FALSE,results='hide',messages=F,error=FALSE}
## this chunk outputs a copy of this script converted to a 'normal' R file with comments
purl("workflow/5_PostFireTrajectoriesAnnual/PostFireTrajectories.Rmd",
     documentation=2,output = "workflow/5_PostFireTrajectoriesAnnual/PostFireTrajectories.R", quiet = TRUE) 
```
